Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2023-12-21 paper.tex
More about anisotropic samples

@@

The main advantage of our algorithm is its parallelizability and ease of
implementation. An equivalent operation for rotation of test triangles is a
rotation of the whole input image which we have implemented in our
CorrelationFunctions.jl package. By applying a rotation to the input first we
made a rest of the algorithm cache-friendly by avoiding random memory access.
\begin{figure*}[tp]
  \centering
  \subfigure[An original image of an anisotropic pore.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso.png}
    \label{fig:aniso-unrot}}
  \hfill
  \subfigure[A rotated image of the pore, so it is elongated along the axis
    $X$.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso-rot.png}
    \label{fig:aniso-rot}}
  \vskip\baselineskip
  \subfigure[$S_3$ function for the original image]{
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png}}
  \hfill
  \subfigure[$S_3$ function for the rotated image. A vector $(0.92, -0.4)$ is a
    direction of elongation of the pore.]{
  \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png}}
  \hfill
  \subfigure[$S_3$ function averaged across 100 random rotations.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png}
    \label{fig:aniso-avg}}
  \caption[]{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. A result on \cref{fig:aniso-avg} is similar
    to the algorithm proposed by Malmir~et~al.}
  \label{fig:aniso}
\end{figure*}

>>>>

\subsection{Analysis of anisotropic media}
When an analized sample is anisotropic, i.e. it is stretched or oriented along
some specific direction, it may be desired to align one of the catheti of the
scanning pattern \cref{fig:pattern} with that direction. In this case the
anisotropicity present in the sample will be featured in a produced correlation
function \textcolor{red}{most well. voobshe eto po-angliyski?}

To detect the direction of anisotropy we use an $N \times N$ (where $N$ is
dimensionality of the sample) covariance matrix for coordinates of points in the
phase of interest:
\begin{equation}
  C_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and decompose $C$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation (by
multiplication of one of the columns by $-1$ if needed, so that
$\det T = 1$. Application of this rotation to the sample makes the direction of
anisotropicity parallel to one of the axes, so anisotropicity can be better
featured in the correlation function \cref{fig:aniso}.
\begin{figure*}[tp]
  \centering
  \subfigure[An original image of an anisotropic pore.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso.png}
    \label{fig:aniso-unrot}}
  \hfill
  \subfigure[A rotated image of the pore, so it is elongated along the axis
    $X$.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso-rot.png}
    \label{fig:aniso-rot}}
  \vskip\baselineskip
  \subfigure[$S_3$ function for the original image]{
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png}}
  \hfill
  \subfigure[$S_3$ function for the rotated image. A vector $(0.92, -0.4)$ is a
    direction of elongation of the pore.]{
  \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png}}
  \hfill
  \subfigure[$S_3$ function averaged across 100 random rotations.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png}
    \label{fig:aniso-avg}}
  \caption[]{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. A result on \cref{fig:aniso-avg} averaged
    across rotations is similar to the algorithm proposed by Malmir~et~al.}
  \label{fig:aniso}
\end{figure*}
The above said holds for every algorithm introduced in the paper.

When a sample is known to be isotropic, the correlation function may be
calculated multiple times, each time applying a random rotation of the sample,
and then averaged across rotations to gather more statistics. This process is
illustrated on \cref{fig:workflow}.
\begin{figure*}[tp]
  \centering
  \includegraphics[width=0.6\linewidth]{images/workflow.png}
  \caption[]{A possible workflow for computation of correlation functions
    (clockwise from top left to bottom right ): 1.~rotating a sample in random
    directions (which is equivalent to rotating the sampling pattern),
    2.~calculating the correlation function for each direction, 3.~avegaring
    results across rotations, 4.~representing the result as a family of 2D
    plots.}
  \label{fig:workflow}
\end{figure*}

@@

The main advantage of our algorithm is its parallelizability and ease of
implementation. An equivalent operation for rotation of test triangles is a
rotation of the whole input image which we have implemented in our
CorrelationFunctions.jl package. By applying a rotation to the input first we
made a rest of the algorithm cache-friendly by avoiding random memory access.
